---
title: 'PHP 2514 Homework #3'
author: "Blain Morin"
date: "March 19, 2018"
output: html_document
---

#Question 1:

Assume X is in units of $10,000.

First, find y-intercept ($\beta_0$):

$$
\begin{align}
\pi &= .27\\
x &= 0\\
logit(.27) &= \ln(\frac{.27}{1-.27}) = \beta_0 + \beta_1X\\
-.9946 &= \beta_0
\end{align}
$$

Next, find the slope ($\beta_1$):

$$
\begin{align}
\pi &= .88 \\
x &= 6 \\
logit(.88) &=\ln(\frac{.88}{1-.88}) = -.9946 +\beta_1 *6\\
.4978 &= \beta_1
\end{align}
$$

Thus, the equation is:

$$
\pi = logit^{-1}(-.9946 + .4978X)
$$

#Question 2

##a.)
```{r, message=FALSE, warning=FALSE}

library(tidyverse)
library(arm)

###create data set
studentid = 1:50
set.seed(50)
grades = rnorm(50, 60, 15)

###use the our logit to predict if a student passed using their midterm grade
prob.pass = c()
for (i in 1:50) {
  
  prob = invlogit(-24 + .4 * grades[i])
  prob.pass[i] = rbinom(1, 1, prob = prob)
  
}

classdata = data.frame(studentid, grades, prob.pass)



###graph our points with the logit function given by the problem
classdata %>%
  ggplot(aes(x = grades, y = prob.pass)) + 
    stat_function(fun = function(x) invlogit(-24 + .4*x), color = "aquamarine3", size = 2) +
    geom_point(size = 3.5, alpha = .4, color = "coral3") + 
    theme_minimal() +
    xlab("Midterm Grade") +
    ylab("Pr(pass)") +
    ggtitle("Passed class vs Midterm Grade")
  
```

##b.)

```{r}

###Add a column with standardized grades
classdata = classdata %>%
  mutate(stand.grade = (grades - mean(grades)) / sd(grades))


###Plot our standardized data with our updated logit curve
###To get our new logit function we multiplied the B0 by 0
###and multiplied B1 by 15
classdata %>%
    ggplot(aes(x = stand.grade, y = prob.pass)) + 
    stat_function(fun = function(x) invlogit((-24*0) + (.4*15)*x), color = "aquamarine3", size = 2) +
    geom_point(size = 3.5, alpha = .4, color = "coral3") + 
    theme_minimal() +
    xlab("Standardized Grade") +
    ylab("Pr(pass)") +
    ggtitle("Pr(Pass) vs Standardized Grade")

```


##c.)

```{r, message = FALSE}
library(sjPlot)
attach(classdata)

###Add random noise column to our data
noise = rnorm(50, 0 , 1)
classdata$noise = noise

###Run model with no noise
model.no.noise = glm(prob.pass ~ stand.grade, family = binomial(link = "logit"))


###Run model with noise
model.noise = glm(prob.pass ~ stand.grade + noise, family = binomial(link = "logit"))


print(c(model.no.noise$deviance, model.noise$deviance))

detach(classdata)

```

The deviance in our noise model is very similar to the deviance in the first model (20.93 vs 20.74. This makes sense theoretically: since the noise is randomly generated, it shouldn't impact the fit of our model (except for some random spurious correlation). Thus, our residuals will not change much. Therefore, the deviance will not change much. 


#Question 3


##a.)

First we examined the histogram for our race categories. 

```{r}

rodents = read.table("rodents.txt")
rodents$rodent2 = as.factor(rodents$rodent2)
rodents$race = as.factor(rodents$race)


hist(as.numeric(rodents$race))
```

We see that race indicators 6 and 7 have a low frequency, so we will bin those together: 

```{r}

###Combine race 6 and 7 into category 6. 

race.edit = c()

for (i in 1:nrow(rodents)) {
  if (rodents$race[i] == 7) {
    race.edit[i] = 6
  } else {
    race.edit[i] = rodents$race[i]
  }
}

rodents$race.edit = as.factor(race.edit)

base.model = glm(formula = rodent2 ~ race.edit,
                 family = binomial(link = "logit"), 
                 data = rodents)

```

```{r, include = FALSE}

summary(base.model)

stargazer::stargazer(base.model, type = "html")

```

<table style="text-align:center"><tr><td colspan="2" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left"></td><td><em>Dependent variable:</em></td></tr>
<tr><td></td><td colspan="1" style="border-bottom: 1px solid black"></td></tr>
<tr><td style="text-align:left"></td><td>rodent2</td></tr>
<tr><td colspan="2" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">race.edit2</td><td>1.536<sup>***</sup></td></tr>
<tr><td style="text-align:left"></td><td>(0.169)</td></tr>
<tr><td style="text-align:left"></td><td></td></tr>
<tr><td style="text-align:left">race.edit3</td><td>1.449<sup>***</sup></td></tr>
<tr><td style="text-align:left"></td><td>(0.214)</td></tr>
<tr><td style="text-align:left"></td><td></td></tr>
<tr><td style="text-align:left">race.edit4</td><td>1.867<sup>***</sup></td></tr>
<tr><td style="text-align:left"></td><td>(0.187)</td></tr>
<tr><td style="text-align:left"></td><td></td></tr>
<tr><td style="text-align:left">race.edit5</td><td>0.400</td></tr>
<tr><td style="text-align:left"></td><td>(0.292)</td></tr>
<tr><td style="text-align:left"></td><td></td></tr>
<tr><td style="text-align:left">race.edit6</td><td>1.364<sup>**</sup></td></tr>
<tr><td style="text-align:left"></td><td>(0.554)</td></tr>
<tr><td style="text-align:left"></td><td></td></tr>
<tr><td style="text-align:left">Constant</td><td>-2.152<sup>***</sup></td></tr>
<tr><td style="text-align:left"></td><td>(0.128)</td></tr>
<tr><td style="text-align:left"></td><td></td></tr>
<tr><td colspan="2" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">Observations</td><td>1,522</td></tr>
<tr><td style="text-align:left">Log Likelihood</td><td>-760.124</td></tr>
<tr><td style="text-align:left">Akaike Inf. Crit.</td><td>1,532.248</td></tr>
<tr><td colspan="2" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left"><em>Note:</em></td><td style="text-align:right"><sup>*</sup>p<0.1; <sup>**</sup>p<0.05; <sup>***</sup>p<0.01</td></tr>
</table>

As you can see in the table above, race 1 if our reference category. Raising e to the beta is the multiplicitive change in the odds of having rodents in the apartment. For example, for race category 2: 

```{r}

exp(1.536)

```

This is interpretted as: 

Compared to race category 1, the odds of having rodents in the apartment is 4.6 times higher for those people in race category 2. 

##b.)

Next, we try to find a good model using other other house-hold level variables. The variables that I'm choosing to examine are:

- Rotted / Broken Windows
- Missing or Worn Flooring
- Cracks in Walls
- Holes in the flooring
- If the building was built before 1947
- If there is a regular exterminator
- Total household income
- Housing type (rent, public housing, owned)

```{r, message=FALSE, results="hide"}

attach(rodents)

b.data = data.frame(
  as.factor(rodent2),
  as.factor(extwin4_2),
  as.factor(extflr5_2),
  as.factor(intcrack2),
  as.factor(inthole2),
  as.factor(old),
  as.factor(race.edit),
  as.factor(regext),
  totincom2
)

detach(rodents)

attach(b.data)

###change income units to thousand
b.data$totincom2 = b.data$totincom2/10000

###remove NAs
b.data = na.omit(b.data)

b.model = glm(as.factor.rodent2. ~ ., family = binomial(link="logit"), data = b.data)

b.step.model = step(b.model, direction = "both")

```

```{r, echo=FALSE}
sjt.glm(b.step.model)
```

The odds ratios for all of the variables are given in the above table. The interpretation of these is similar to part(a). For example, having a hole in the floor of the apartment means that there is 2.88 times the odds of having rodents than in an apartment with no holes in the floor. The interpretation is similar for income: a household earning 1000 dollars more than another similar household has .95 times the odds of having rodents. 

Overall, most of the variables in our model are significant at the 5 percent level. The signs of the household factors are consistent with what one would expect (ie holes in the floor is associated with higher odds of having rats). The variables that the model dropped were having a regular exterminator and having a crack in the exterior windows. 

#Question 4

```{r, message=FALSE}

dhaka = read_csv("dhaka.csv")

library(nnet)
library(sjPlot)

###Select clinical only columns
clinicaldata = dhaka[, 1:13]

###Run regression for only clinical predictors
model.clinical = multinom(data = clinicaldata,
                          formula = dehyd ~ .)

###Run regression for all predictors
model.all = multinom(data = clinicaldata,
                     formula = dehyd ~ .)

summary(model.all)
```
<table style="text-align:center"><tr><td colspan="3" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left"></td><td colspan="2"><em>Dependent variable:</em></td></tr>
<tr><td></td><td colspan="2" style="border-bottom: 1px solid black"></td></tr>
<tr><td style="text-align:left"></td><td>1</td><td>2</td></tr>
<tr><td style="text-align:left"></td><td>(1)</td><td>(2)</td></tr>
<tr><td colspan="3" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">capref</td><td>11.605<sup>***</sup></td><td>-5.306<sup>***</sup></td></tr>
<tr><td style="text-align:left"></td><td>(0.00000)</td><td>(0.000)</td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td style="text-align:left">extrem</td><td>0.158</td><td>0.628</td></tr>
<tr><td style="text-align:left"></td><td>(0.869)</td><td>(0.943)</td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td style="text-align:left">eyes</td><td>0.262</td><td>0.867<sup>*</sup></td></tr>
<tr><td style="text-align:left"></td><td>(0.310)</td><td>(0.474)</td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td style="text-align:left">genapp</td><td>0.600<sup>***</sup></td><td>0.887<sup>***</sup></td></tr>
<tr><td style="text-align:left"></td><td>(0.208)</td><td>(0.283)</td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td style="text-align:left">heart</td><td>0.597<sup>*</sup></td><td>0.329</td></tr>
<tr><td style="text-align:left"></td><td>(0.330)</td><td>(0.449)</td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td style="text-align:left">mucous</td><td>-0.180</td><td>0.325</td></tr>
<tr><td style="text-align:left"></td><td>(0.297)</td><td>(0.540)</td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td style="text-align:left">pulse</td><td>0.673<sup>**</sup></td><td>0.692<sup>*</sup></td></tr>
<tr><td style="text-align:left"></td><td>(0.330)</td><td>(0.393)</td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td style="text-align:left">resp</td><td>0.489</td><td>0.241</td></tr>
<tr><td style="text-align:left"></td><td>(0.368)</td><td>(0.479)</td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td style="text-align:left">skin</td><td>0.873<sup>***</sup></td><td>0.853<sup>**</sup></td></tr>
<tr><td style="text-align:left"></td><td>(0.292)</td><td>(0.417)</td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td style="text-align:left">tears</td><td>0.083</td><td>0.455</td></tr>
<tr><td style="text-align:left"></td><td>(0.258)</td><td>(0.356)</td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td style="text-align:left">thirst</td><td>-0.489</td><td>0.238</td></tr>
<tr><td style="text-align:left"></td><td>(0.509)</td><td>(0.613)</td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td style="text-align:left">urine</td><td>0.073</td><td>0.196</td></tr>
<tr><td style="text-align:left"></td><td>(0.211)</td><td>(0.330)</td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td style="text-align:left">Constant</td><td>-0.796</td><td>-4.547<sup>***</sup></td></tr>
<tr><td style="text-align:left"></td><td>(0.586)</td><td>(0.855)</td></tr>
<tr><td style="text-align:left"></td><td></td><td></td></tr>
<tr><td colspan="3" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left">Akaike Inf. Crit.</td><td>645.875</td><td>645.875</td></tr>
<tr><td colspan="3" style="border-bottom: 1px solid black"></td></tr><tr><td style="text-align:left"><em>Note:</em></td><td colspan="2" style="text-align:right"><sup>*</sup>p<0.1; <sup>**</sup>p<0.05; <sup>***</sup>p<0.01</td></tr>
</table>
